Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-07-30 literature/paper_draft.qmd
cleanup

-   **Problem Statement:**

@@

    -   A literature review of forecast reconciliation and covariance estimation.
    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
    -   An experimental design using both synthetic and real hierarchical time series.
    -   Empirical results and discussion.
    -   Conclusions and suggestions for future work.

# Literature Review

@@

<!-- Wickramasuriya et al. (2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.  However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).  Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U). -->

<!-- Empirical findings to date   -->

<!-- Athanasopoulos et al. (2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.  Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators. -->

@@

-   Discuss how covariance estimation affects MinT performance.
-   Identify research gaps.

@@

However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage. Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions. We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$. The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen. We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.

@@

For each series, $T = 54$ and $304$ observations are generated. The first 50 and 300 observations are used for training, and the last 4 observations are used for testing. The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa]. We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators. These include using the unbiased sample covariance matrix - mint_sample, the shrinkage estimator - mint_shr, and the NOVELIST estimator - mint_n. The Monte Carlo simulation is repeated $M = 200$ times, in which the parameters for data generating process is fixed.

@@

## Exploring Effects of Hierarchy's Aggregation

@@

- Finding: Complex aggregation MinT_n performs well

## h-step-ahead forecast error metric in NOVELIST Cross-validation

Instead of forecasting t+1 for each rolling window iteration, I forecast t+1 to t+h for each iteration and compute the MSE for each threshold value delta.

Results barely show any difference with the original method.

```{r}
path <- "./literature/lib/hstep_CV_Novelist/"

```

@@

Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which splits into the seven states and territories. Each state is further sub-divided into tourism “zones,” which in turn break down into 77 “regions.” Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.

<!-- Table to illustrate the hierarchy -->

We quantify tourism demand via “visitor nights,” the total number of nights spent by Australians away from home. The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourismau-jk].

To assess forecasting performance, we adopt a rolling-window cross-validation scheme. Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@fabletools-fa]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models. These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n). We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018. In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.

@@

@@

Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise. This contradicts with the canonical results from @Wickramasuriya2019-xq and others, prompting further investigation. Applying STL decomposition to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism. The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.

@@

# Discussion and Conclusion

@@

-   Limitations of the research

>>>>

<!--
-   **Problem Statement:**

@@

    -   A literature review of forecast reconciliation and covariance estimation.
    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
    -   An experimental design using both synthetic and real hierarchical time series.
    -   Empirical results and discussion.
    -   Conclusions and suggestions for future work.
-->

<!--
# Literature Review

@@

Wickramasuriya et al. (2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.  However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).  Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).

Empirical findings to date

Athanasopoulos et al. (2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.  Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.

@@

-   Discuss how covariance estimation affects MinT performance.
-   Identify research gaps.
-->

@@

However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage. Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions. We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$. The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen. We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.

@@

For each series, $T = 54$ and $304$ observations are generated. The first 50 and 300 observations are used for training, and the last 4 observations are used for testing. The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@O-Hara-Wild2024-we]. We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators. These include using the unbiased sample covariance matrix - mint_sample, the shrinkage estimator - mint_shr, and the NOVELIST estimator - mint_n. The Monte Carlo simulation is repeated $M = 200$ times, in which the parameters for data generating process is fixed.

@@

<!--
## Exploring Effects of Hierarchy's Aggregation

@@

- Finding: Complex aggregation MinT_n performs well
-->

@@

Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which splits into the seven states and territories. Each state is further sub-divided into tourism zones, which in turn break down into 77 regions. Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.

- A table/figure to illustrate the hierarchy

We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home. The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].

@@

To assess forecasting performance, we adopt a rolling-window cross-validation scheme. Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models. These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n). We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018. In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.

@@

Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise. This contradicts with the canonical results from @Wickramasuriya2019-xq, prompting further investigation. Applying STL decomposition by @Bandara2022-og to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism. The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.

@@

# Timeline & Milestones

<!--
# Discussion and Conclusion

@@

-   Limitations of the research
-->
